### 2020-02-17 Article/LaTeX_DL_468198_240419/calibration_parameters.tex
Calibration Updated

\documentclass{standalone}
\usepackage{tikz}
%\usetikzlibrary{patterns}
\usetikzlibrary{calc}
%\usepackage{graphicx}

\tikzset{
    right angle quadrant/.code={
        \pgfmathsetmacro\quadranta{{1,1,-1,-1}[#1-1]}     % Arrays for selecting quadrant
        \pgfmathsetmacro\quadrantb{{1,-1,-1,1}[#1-1]}},
    right angle quadrant=1, % Make sure it is set, even if not called explicitly
    right angle length/.code={\def\rightanglelength{#1}},   % Length of symbol
    right angle length=2ex, % Make sure it is set...
    right angle symbol/.style n args={3}{
        insert path={
            let \p0 = ($(#1)!(#3)!(#2)$) in     % Intersection
                let \p1 = ($(\p0)!\quadranta*\rightanglelength!(#3)$), % Point on base line
                \p2 = ($(\p0)!\quadrantb*\rightanglelength!(#2)$) in % Point on perpendicular line
                let \p3 = ($(\p1)+(\p2)-(\p0)$) in  % Corner point of symbol
            (\p1) -- (\p3) -- (\p2)
        }
    }
}

\begin{document}
\def\s{9}
\def\c{3}
\begin{tikzpicture}[font=\large]
\coordinate (CL) at (0,\c);
\coordinate (CR) at (\c,0);
\coordinate (CC) at (\c/2,\c/2);
\coordinate (P) at (\c,\c);
\coordinate (HL) at (0, \s);
\coordinate (HR) at (15,\s);
\coordinate(HC) at (\s,\s);
\coordinate (N) at (\c,\s);
\coordinate (TLF) at (0.75*\c, 0.5*\c);

\draw(CL)--(CR);
\draw(CC)--(P) node [midway, above left] {$L_f$};
\node[circle,fill,inner sep=1pt, label={below left:$(x_p^0, y_p^0)$}] at (CC) {};
\node[circle,fill,inner sep=1pt, label={right:$P$}] at (P) {};
\draw(P)--(HC) node [midway, above left] {$L_m$};
\draw[thick](HL)--(HR);
\draw[dashed](N)--(P);
\node[label={[align=left]below:$ax+by+cz+d = 0$\\$d = - c L_m$ }] at (HR) {};

\draw[thick,->](N)--(\c,\s-1);
\node[label={right:$\underline{\hat{n}} = (a, b, c)$}] at (\c,\s-0.5) {};

\node[label={right:$\frac{1}{f}= \frac{1}{L_m} + \frac{1}{L_f}$}] at (TLF)  {};

\draw [right angle quadrant=2,right angle symbol={HL}{HR}{P}];
\draw [right angle quadrant=1,right angle symbol={CL}{CR}{P}];
\end{tikzpicture}

\end{document}